% !TEX program = pdflatex
\input{jmlr_coverletter_preamble.tex}

\begin{document}

\begin{flushright}
Tatsuki Onishi\\
Data Science and AI Innovation Research Promotion Center\\
Shiga University, Japan\\
\texttt{bougtoir@gmail.com}\\[6pt]
\today
\end{flushright}

Journal of Machine Learning Research

\bigskip

Dear Editors:

I am writing to submit my manuscript ``Spectral Causality: Causal
Direction Estimation via Magnetic Laplacians and Hodge Decomposition''
to the \emph{Journal of Machine Learning Research}.

The manuscript introduces \emph{spectral causality}, a principled framework
that estimates causal directions by exploiting the spectral structure of the
magnetic Laplacian.  I define a Directional Predictability Index (DPI) that
resolves circularity in earlier utility-based formulations, connect the
framework to Hodge decomposition to separate DAG-compatible and feedback
components of causal flow, and establish scale-invariance and
phase-transition results.  The approach is validated on synthetic DAGs, the
Sachs protein signaling network, and the UCI Heart Disease dataset, with
benchmarks against DirectLiNGAM, PC, GES, and NOTEARS.

This paper has not been published previously in any journal or conference
proceedings.  A preprint may be made available on arXiv concurrent with this
submission.

I suggest the following action editors and referees for this submission.

Action Editors:
\begin{itemize}
  \item Bryon Aragam, University of Chicago (bragon@uchicago.edu) --- causality, graphical models
  \item Elias Bareinboim, Columbia University (eb@cs.columbia.edu) --- causal inference, generalizability
  \item Silvia Chiappa, DeepMind (silviac@google.com) --- causal inference, variational inference
  \item Kenji Fukumizu, The Institute of Statistical Mathematics, Japan (fukumizu@ism.ac.jp) --- kernel methods, dimension reduction
  \item Mladen Kolar, University of Southern California (mkolar@usc.edu) --- graphical models, high-dimensional statistics
\end{itemize}

Reviewers:
\begin{itemize}
  \item Joris Mooij, University of Amsterdam (j.m.mooij@uva.nl) --- causal discovery, structural causal models
  \item Dominik Janzing, Amazon T\"ubingen (janzind@amazon.com) --- causal inference, information-theoretic methods
  \item Biwei Huang, UC San Diego (bih007@ucsd.edu) --- causal discovery, latent variable models
  \item Murat Kocaoglu, Purdue University (mkocaoglu@purdue.edu) --- causal inference, graphical models
  \item Frederick Eberhardt, California Institute of Technology (fde@caltech.edu) --- causal inference, Bayesian networks
\end{itemize}

This submission has the following keywords: causal discovery, magnetic
Laplacian, Hodge decomposition, spectral graph theory, directed graphs.

As the corresponding (and sole) author, I confirm that I have no conflict
of interest with the action editors and referees suggested above.

\medskip

\textbf{Disclosure of funding and competing interests.}
No external funding was received for this work.
The author declares no competing interests.

\bigskip

Sincerely,

\medskip

Tatsuki Onishi\\
Data Science and AI Innovation Research Promotion Center\\
Shiga University, Japan

\end{document}
